flowCore: data structures package for flow cytometry data

### 2007-04-16 inst/doc/HowTo-flowCore.Rnw
Fix vignette and man page
git-svn-id: https://hedgehog.fhcrc.org/bioconductor/trunk/madman/Rpacks/flowCore@23904 bc3139a8-67e5-0310-9ffc-ced21a209358

\title{flowCore: data structures pacakge for flow cytometry data}
\author{N. Le Meur F. Hahne B. Ellis P. Haaland}

@@

<<echo=false,results=hide>>=
library(flowCore)
@

@@

<<echo=true,results=verbatim>>=
file.name <- system.file("extdata","0877408774.B08", package="flowCore")
x = read.FCS(file.name,transformation=FALSE)
summary(x)
@

@@

<<echo=true,results=verbatim>>=
keyword(x,c("$P1E","$P2E","$P3E","$P4E"))
@

@@

<<echo=true,results=verbatim>>=
summary(read.FCS(file.name))
@

@@

<<echo=true,results=verbatim>>=
summary(read.FCS(file.name,transformation="scale"))
@

@@

<<echo=true,results=verbatim>>=
read.FCS(file.name,alter.names=TRUE)
@

@@

<<echo=true,results=verbatim>>=
x = read.FCS(file.name,column.pattern="-H")
x
@

@@

<<fig=true,echo=true>>=
plot(x)
@

@@

<<fig=true,echo=true>>=
plot(x,c("FL1-H","FL2-H"))
@

@@

<<fig=true,echo=true>>=
plot(x,"FL1-H",breaks=256)
@

@@

<<echo=true,results=verbatim>>=
frames = lapply(dir(system.file("extdata","compdata","data",package="flowCore"),full.names=TRUE),read.FCS)
as(frames,"flowSet")
@

@@

<<echo=true,results=verbatim>>=
names(frames) = sapply(frames,keyword,"SAMPLE ID")
fs = as(frames,"flowSet")
fs
@

@@

<<echo=true,results=verbatim>>=
phenoData(fs)$Filename = fsApply(fs,keyword,"$FIL")
pData(phenoData(fs))
@

@@

<<echo=true,results=verbatim>>=
read.flowSet(path=system.file("extdata","compdata","data",package="flowCore"))
@

@@

<<echo=true,results=verbatim>>=
fs = read.flowSet(path=system.file("extdata","compdata","data",package="flowCore"),name.keyword="SAMPLE ID",phenoData=list(name="SAMPLE ID",Filename="$FIL"))
fs
pData(phenoData(fs))
@

@@

<<echo=true,results=verbatim>>=
fsApply(fs,each_col,median)
@

@@

<<echo=true,results=verbatim>>=
fsApply(fs,function(x) apply(x,2,median),use.exprs=TRUE)
@

@@

<<fig=true,echo=true>>=
plot(transform(fs[[1]],`FL1-H`=log(`FL1-H`),`FL2-H`=log(`FL2-H`)),c("FL1-H","FL2-H"))
@

@@

<<fig=true,echo=true>>=
plot(transform(fs[[1]],log.FL1.H=log(`FL1-H`),log.FL2.H=log(`FL2-H`)),c("log.FL1.H","log.FL2.H"))
@

@@

\begin{description}
\item[truncateTransform]{ $y = \left\{\begin{array}{ll}
a & x < a \\
x & x \geq a \\
\end{array}\right.$ }
\item[scaleTransform]{$f(x) = \frac{x-a}{b-a}$}
\item[linearTransform]{ $f(x) = a+bx$ }
\item[quadraticTransform]{ $f(x) = ax^2 + bx + c$ }
\item[lnTransform]{$ f(x) = \log\left(x\right)\frac{r}{d} $}
\item[logTransform]{$ f(x) = \log_b\left(x\right)\frac{r}{d} $}
\item[biexponentialTransform]{ $f^{-1}(x) = ae^{bx}-ce^{dx}+f$}
\item[logicleTransform]{ A special form of the biexponential transform with parameters selected by the data.}
\item[arcsinhTransform]{ $f(x) = \asinh\left(a+bx\right)+c$}
\end{description}

@@

<<echo=true,results=verbatim>>=
aTrans = truncateTransform("truncate at 1",a=1)
aTrans
@
which we can then use on the parameter of interest in the usual way
<<echo=true,results=verbatim>>=
transform(fs,`FL1-H`=aTrans(`FL1-H`))
@

@@

<<echo=true,results=verbatim>>=
rectGate = rectangleGate(filterId="Fluorescence Region","FL1-H"=c(50,100),"FL2-H"=c(50,100))
@

@@

<<echo=true,results=verbatim>>=
summary(result)
summary(result)$n
summary(result)$true
summary(result)$p
@

@@

<<echo=true,results=verbatim>>=
summary(filter(fs[[1]],kmeansFilter("myKMeans","FSC-H"=c("Low","Medium","High"))))
@

@@

<<echo=true,results=verbatim>>=
morphGate = norm2Filter(filterId="MorphologyGate","FSC-H","SSC-H",scale=2)
smaller = Subset(fs,morphGate)
fs[[1]]
smaller[[1]]
@

@@

<<echo=true,results=verbatim>>=
split(smaller[[1]],kmeansFilter("myKMeans","FSC-H"=c("Low","Medium","High")))
@

@@

<<echo=true,results=verbatim>>=
split(smaller,kmeansFilter("myKMeans","FSC-H"=c("Low","Medium","High")))
@

@@

Of course, most filtering operations consist of more than one gate. To combine gates and filters we use the standard R Boolean operators: \&, \| and ! to construct an intersection, union and complement respectively:

<<echo=true,results=verbatim>>=

@@

<<echo=true,results=verbatim>>=
summary(filter(smaller[[1]],rectGate %&% morphGate))
@

@@

<<echo=true,results=verbatim>>=
tFilter = transform("FL1-H"=log,"FL2-H"=log)
tFilter
@

@@

<<echo=true,results=verbatim>>=
rect2 = rectangleGate(filterId="Another Rect","FL1-H"=c(1,2),"FL2-H"=c(2,3)) %on% tFilter
rect2
@

@@

<<fig=true,echo=true>>=
plot(tFilter %on% smaller[[1]],c("FL1-H","FL2-H"))
@

>>>>

\title{flowCore: data structures package for flow cytometry data}
\author{N. Le Meur F. Hahne B. Ellis P. Haaland}

@@

<<loadPackage, echo=false,results=hide>>=
library(flowCore)
@

@@

<<ReadFiles, echo=true, results=verbatim>>=
file.name <- system.file("extdata","0877408774.B08", package="flowCore")
x <- read.FCS(file.name, transformation=FALSE)
summary(x)
@

@@

<<SearchKeywords, echo=true,results=verbatim>>=
keyword(x,c("$P1E", "$P2E", "$P3E", "$P4E"))
@

@@

<<PrintSummary, echo=true, results=verbatim>>=
summary(read.FCS(file.name))
@

@@

<<PrintSummary2, echo=true,results=verbatim>>=
summary(read.FCS(file.name,transformation="scale"))
@

@@

<<ReadFiles2, echo=true,results=verbatim>>=
read.FCS(file.name,alter.names=TRUE)
@

@@

<<RedFiles3, echo=true, results=verbatim>>=
x <- read.FCS(file.name,column.pattern="-H")
x
@

@@

<<Plot1, fig=true, echo=true>>=
plot(x)
@

@@

<<Plot2, fig=true, echo=true>>=
plot(x,c("FL1-H", "FL2-H"))
@

@@

<<plot3, fig=true, echo=true>>=
plot(x, "FL1-H", breaks=256)
@

@@

<<Frames, echo=true, results=verbatim>>=
frames <- lapply(dir(system.file("extdata", "compdata", "data", package="flowCore"), full.names=TRUE), read.FCS)
as(frames, "flowSet")
@

@@

<<Frames, echo=true,results=verbatim>>=
names(frames) <- sapply(frames, keyword, "SAMPLE ID")
fs <- as(frames, "flowSet")
fs
@

@@

<<metaData, echo=true, results=verbatim>>=
phenoData(fs)$Filename <- fsApply(fs,keyword, "$FIL")
pData(phenoData(fs))
@

@@

<<ReadFlowSet, echo=true,results=verbatim>>=
read.flowSet(path = system.file("extdata", "compdata", "data", package="flowCore"))
@

@@

<<ReadFowSet2, echo=true,results=verbatim>>=
fs <- read.flowSet(path=system.file("extdata", "compdata", "data", package="flowCore"), name.keyword="SAMPLE ID", phenoData=list(name="SAMPLE ID", Filename="$FIL"))
fs
pData(phenoData(fs))
@

@@

<<fsApply1, echo=true, results=verbatim>>=
fsApply(fs, each_col, median)
@

@@

<<fsApply2, echo=true,results=verbatim>>=
fsApply(fs,function(x) apply(x, 2, median), use.exprs=TRUE)
@

@@

<<Transfo1, fig=true, echo=true>>=
plot(transform(fs[[1]], `FL1-H`=log(`FL1-H`), `FL2-H`=log(`FL2-H`)), c("FL1-H","FL2-H"))
@

@@

<<Transfo2, fig=true, prefix=FALSE, include=FALSE, echo=true>>=
plot(transform(fs[[1]], log.FL1.H=log(`FL1-H`), log.FL2.H=log(`FL2-H`)), c("log.FL1.H", "log.FL2.H"))
@

\begin{figure}
\centering
\includegraphic{Transfo2}
\end{figure}

@@

\begin{description}
\item[truncateTransform]{ $y = \left\{\begin{array}{ll}
a & x < a \\
x & x \geq a \\
\end{array}\right.$ }
\item[scaleTransform]{$ f(x) = \frac{x-a}{b-a}$}
\item[linearTransform]{ $ f(x) = a+bx$ }
\item[quadraticTransform]{ $ f(x) = ax^2 + bx + c$ }
\item[lnTransform]{$ f(x) = \log\left(x\right)\frac{r}{d} $}
\item[logTransform]{$ f(x) = \log_b\left(x\right)\frac{r}{d} $}
\item[biexponentialTransform]{ $ f^{-1}(x) = ae^{bx}-ce^{dx}+f$}
\item[logicleTransform]{ A special form of the biexponential transform with parameters selected by the data.}
\item[arcsinhTransform]{ $ f(x) = asinh\left(a+bx\right)+c$}
\end{description}

@@

<<Transfo3, echo=true>>=
aTrans <- truncateTransform("truncate at 1", a=1)
aTrans
@
which we can then use on the parameter of interest in the usual way
<<Transfo4, echo=true,results=verbatim>>=
transform(fs,`FL1-H`=aTrans(`FL1-H`))
@

@@

<<rectGate, echo=true, results=verbatim>>=
rectGate <- rectangleGate(filterId="Fluorescence Region", "FL1-H"=c(50, 100), "FL2-H"=c(50, 100))
@

@@

<<Summary3, echo=true, results=verbatim>>=
summary(result)
summary(result)$n
summary(result)$true
summary(result)$p
@

@@

<<SummarFilter, echo=true, results=verbatim>>=
summary(filter(fs[[1]], kmeansFilter("myKMeans","FSC-H"=c("Low", "Medium", "High"))))
@

@@

<<Norm2Filter, echo=true, results=verbatim>>=
morphGate <- norm2Filter(filterId="MorphologyGate", "FSC-H", "SSC-H", scale=2)
smaller <- Subset(fs, morphGate)
fs[[1]]
smaller[[1]]
@

@@

<<Split, echo=true, results=verbatim>>=
split(smaller[[1]], kmeansFilter("myKMeans","FSC-H"=c("Low","Medium","High")))
@

@@

<<Split2, echo=true, results=verbatim>>=
split(smaller, kmeansFilter("myKMeans", "FSC-H"=c("Low", "Medium", "High")))
@

@@

Of course, most filtering operations consist of more than one gate. To combine gates and filters we use the standard R Boolean operators: \&, | and ! to construct an intersection, union and complement respectively:

<<CombineFilter, echo=true, results=verbatim>>=

@@

<<Summary5, echo=true, results=verbatim>>=
summary(filter(smaller[[1]],rectGate %&% morphGate))
@

@@

<<Transfo5, echo=true,results=verbatim>>=
tFilter <- transform("FL1-H"=log,"FL2-H"=log)
tFilter
@

@@

<<TectGate3, echo=true, results=verbatim>>=
rect2 <- rectangleGate(filterId="Another Rect", "FL1-H"=c(1,2), "FL2-H"=c(2,3)) %on% tFilter
rect2
@

@@

<<Plot6, fig=true,echo=true>>=
plot(tFilter %on% smaller[[1]],c("FL1-H","FL2-H"))
@
